---
title: "Jonas"
author: "Jonas"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library("readxl")
library(glmnet)
library(dplyr)
library(tidyr)
```

## R Markdown

This dataset we used for our analysis is the Food Environment Atlas, taken from the USDA Economic Research Service. This datatset contains information on a number of food related environmental factors such as proximity to stores, food prices, food assistance programs, local food availability. It also includes data on factors that may be related to these such as obesity rate, diabetes rate, and physical activity. This data is provided on the country level across the United States. The purpose of the food environment atlas is to foster research in these areas, and give an overview of food access and health.

The data for the Food Environment Atlas is largely taken from CDC estimates. The CDC uses data from the Behavioral Risk Factor Surveillance System which is under the US Census Bureau to make estimates. 

We set out to answer the question:  What are the best predictors of adult diabetes rate in the US?

Diabetes rate in this dataset is defined as the rate (percent) of persons over 20 years old with a BMI over 30 within each county. We compared this rate at the county level between other factors in this dataset such as proximity to grocery stores, food insecurity, socioeconomic factors and more to determine the best predictors.


```{r import data}
path <- "data/FoodEnvironmentAtlas.xls"
df_all <- path %>%
  excel_sheets() %>%
  set_names() %>%
  map(read_excel, path = path)
```

## Including Plots

You can also embed plots, for example:

```{r EDA}
glimpse(df_all[9])
df_access <- df_all[[5]]
df_stores <- df_all[[6]]
df_restaurants <- df_all[[7]]
df_assistance <- df_all[[8]]
df_insecurity <- df_all[[9]]
df_taxes <- df_all[[10]]
df_local <- df_all[[11]]
df_health <- df_all[[12]]
df_socioeconomic <- df_all[[13]]
```


PCT_DIABETES_ADULTS08
"Diabetes rate 2008"

PCT_DIABETES_ADULTS13
"Diabetes rate 2013"

PCT_SNAP16
"SNAP participants (% pop), 2016*"

PCT_SNAP12
"SNAP participants (% pop), 2012*"

VLFOODSEC_13_15
"Household very low food security (%, three-year average), 2013-15*"

```{r}
df_assist_insec <-
  inner_join(
    df_assistance,
    df_insecurity,
    by = c("FIPS", "State", "County")
  )


```



```{r}
df_assist_insec %>%
  ggplot(aes(PCT_SNAP12, VLFOODSEC_13_15)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}
df_assist_insec %>%
  ggplot(aes(PCT_SNAP12, VLFOODSEC_13_15)) +
  geom_point(aes(color = State))

```

```{r}
df_health_socio <-
  inner_join(
    df_health,
    df_socioeconomic,
    by = c("FIPS", "State", "County")
  )
```


```{r}

df_health_1  <- df_health_socio %>%
  na.omit() %>%
  select(-c('FIPS', "State", "County"))

x_vars <- model.matrix(PCT_DIABETES_ADULTS08~. , df_health_1)[,-1]
y_var <- df_health_1$PCT_DIABETES_ADULTS08
lambda_seq <- 10^seq(2, -2, by = -.1)
grid = 10^seq(10, -2, length = 100)
```



```{r}
set.seed(86)
train = sample(1:nrow(x_vars), nrow(x_vars)/2)
x_test = (-train)
y_test = y_var[x_test]

```

```{r}
lasso_mod = glmnet(x_vars[train,], 
                   y_var[train], 
                   alpha = 1, 
                   lambda = grid) # Fit lasso model on training data

plot(lasso_mod)    # Draw plot of coefficients
lasso_mod
```



```{r}
set.seed(1)
cv.out = cv.glmnet(x_vars[train,], y_var[train],  alpha = 1) # Fit lasso model on training data
plot(cv.out) # Draw plot of training MSE as a function of lambda
bestlam = cv.out$lambda.min # Select lamda that minimizes training MSE

```

```{r}
lasso_best <- glmnet(x_vars[train,], y_var[train], alpha = 1, lambda = bestlam)
pred <- predict(lasso_best, s = bestlam, newx = x_vars[x_test,])
```

```{r}
final <- cbind(y_test, pred)
# Checking the first six obs
head(final)
```




```{r}
# Plot the top 10 coefficients
coef(lasso_best)
plot(sort(coef(lasso_best), decreasing = TRUE)[1:10])
```



```{r}


```


```{r}
# Plot of test data vs predicted with using all nonzero predictors with error bars

final <- cbind(y_var[x_test], pred)



```







```{r}
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),
    strip.text.x = element_text(size = 12),
    strip.text.y = element_text(size = 12),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90"),
    aspect.ratio = 4 / 4,
    plot.margin = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    plot.title = element_text(size = 18),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 16),
    plot.caption = element_text(size = 12)
  )
}
```



